---
title: "EDA Frequncy of Water peril claims in CovA. Distribution"
output:
  html_document:
    df_print: paged
---

```{r}
dataset <- read.csv("C:\\Kate\\Research\\Property\\Data\\property_water_claims_non_cat_fs_v5.csv", header=TRUE)
```



```{r}
library(vcd)
library(ggplot2)
library(fitdistrplus)
```

```{r}
colnames(dataset)
```

## Quick Overview





```{r}
str(dataset)
```


```{r}
summary(dataset)
```

## Number of Claims - Poisson Distribution

We can use goodfit() from vcd package. 
H0: The process is a Poisson process.

```{r}
gf = goodfit(dataset$cova_ic_nc_water,type= "poisson",method= "ML")
summary(gf)
```

Plot

```{r}
plot(gf,main="Number of Water Claims data vs Poisson distribution")
```

P-Value

```{r}
gf.summary = capture.output(summary(gf))[[5]]
pvalue = unlist(strsplit(gf.summary, split = " "))
pvalue = as.numeric(pvalue[length(pvalue)]) 
pvalue
```

P-Value is less then 0.05 and we can reject H0


It is not a Poisson distribution

## Number of Claims - Negative Binomial Distribution


```{r}
gf = goodfit(dataset$cova_ic_nc_water,type= "nbinomial",method= "ML")
summary(gf)
```

```{r}
plot(gf,main="Number of Water Claims data vs Negative Binomial distribution")
```


```{r}
gf.summary = capture.output(summary(gf))[[5]]
pvalue = unlist(strsplit(gf.summary, split = " "))
pvalue = as.numeric(pvalue[length(pvalue)]) 
pvalue
```

P-Value is larger then 0.05 and we can not reject H0 
The distribution is closer to negative binomial

##Incurred Losses

```{r}
data <- read.csv("C:\\Kate\\Research\\Property\\Data\\property_water_claims_non_cat_fs_v5.csv", header=TRUE)
```



Positive Losses only
```{r}
dataset <- data[data$cova_il_nc_water>=100,]
```



And removing long tail of large losses which is better to Gamma distribution but worse for normal log of losses
```{r}
dataset <- dataset[dataset$cova_il_nc_water<quantile(dataset$cova_il_nc_water, 0.95),]
```



```{r}
ggplot(dataset, aes(x = .data[['cova_il_nc_water']])) +
  geom_histogram(bins=100) +
  labs(x = 'cova_il_nc_water', y = 'Count', title = paste("Histogram of", 'cova_il_nc_water'))+
  theme_light()

ggplot(dataset, aes(x = .data[['cova_il_nc_water']])) +
  geom_density() + 
  labs(x = 'cova_il_nc_water', y = 'Density', title = paste("Density of", 'cova_il_nc_water'))

  vec <- dataset$cova_il_nc_water
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]

ggplot(dataset, aes(sample = .data[['cova_il_nc_water']], col='red')) + 
  stat_qq() + 
  geom_abline(slope = slope, intercept = int) +
  theme(legend.position = "none") +
  labs(y = 'cova_il_nc_water', title = paste("QQ Plot of", 'cova_il_nc_water'))

ggplot(dataset, aes(x = .data[['log_cova_il_nc_water']])) +
  geom_histogram(bins=100) +
  labs(x = 'log(cova_il_nc_water)', y = 'Count', title = paste("Histogram of log ", 'cova_il_nc_water'))+
  theme_light()

ggplot(dataset, aes(x = .data[['log_cova_il_nc_water']])) +
  geom_density() + 
  labs(x = 'log(cova_il_nc_water)', y = 'Density', title = 'Density of log(cova_il_nc_water)')

  vec <- dataset$log_cova_il_nc_water
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]

ggplot(dataset, aes(sample = .data[['log_cova_il_nc_water']], col='red')) + 
  stat_qq() + 
  geom_abline(slope = slope, intercept = int) +
  theme(legend.position = "none") +
  labs(y = 'log(cova_il_nc_water)', title = 'QQ Plot of log(cova_il_nc_water)')
```

### Incurred Losses - Gamma Distribution

https://stackoverflow.com/questions/45536234/how-would-you-fit-a-gamma-distribution-to-a-data-in-r
https://www.r-bloggers.com/goodness-of-fit-test-in-r/






```{r}
plot(dataset$cova_il_nc_water, pch=20)
```

```{r}
plotdist(dataset$cova_il_nc_water, histo = TRUE, demp = TRUE)
```

```{r}
descdist(dataset$cova_il_nc_water, discrete=FALSE, boot=500)
```
Scaling is needed to preven "Error in fitdist(dataset$cova_il_nc_water/10, "gamma", method = "mle") : the function mle failed to estimate the parameters, with the error code 100" 

https://stackoverflow.com/questions/53557022/error-code-100-fitting-exp-distribution-using-fitdist-in-r

It looks like some numerical stability problem with the underlying algorithm. It hits something indistinguishable from infinity.

```{r}
fit_w  <- fitdist(dataset$cova_il_nc_water/100, "weibull", method="mle")
fit_g  <- fitdist(dataset$cova_il_nc_water/100, "gamma", method="mle")
fit_ln <- fitdist(dataset$cova_il_nc_water/100, "lnorm", method="mle")
```
```{r}
summary(fit_w)
```
```{r}
summary(fit_g)
```
```{r}
summary(fit_ln)
```

```{r}
plot.legend <- c("Weibull", "gamma", "lognormal")
denscomp(list(fit_w,  fit_g, fit_ln), legendtext = plot.legend)
cdfcomp (list(fit_w,  fit_g, fit_ln), legendtext = plot.legend)
qqcomp  (list(fit_w,  fit_g, fit_ln), legendtext = plot.legend)
ppcomp  (list(fit_w,  fit_g, fit_ln), legendtext = plot.legend)
```
Gamma and weibull are close to each other. The observed distribution is not good for any. Well, the best from worse is Gamma


#### Kolmogorov-Smirnov does not work
```{r}
m <- mean(dataset$cova_il_nc_water)
v <- var(dataset$cova_il_nc_water)

print(m)
print(v)

scale <- v/m
shape <- m*m/v

print(shape)
print(1/scale)
```

Shape = 1.024066
Scale = 35519.68

Kolmogorov-Smirnov is simple nonparametric test for one dimensional probability distribution. Same as Cramer von Mises test, it compares empirical distribution with reference probability. 


```{r}
num_of_samples = 10000
y <- rgamma(num_of_samples, shape = shape, scale = 1/scale)
ks.test(dataset$cova_il_nc_water, y)
```

### Log of Incurred Losses - Normal Distribution


```{r}
dataset <- data[data$cova_il_nc_water>100,]
```

```{r}
ggplot(dataset, aes(x = .data[['cova_il_nc_water']])) +
  geom_histogram(bins=100) +
  labs(x = 'cova_il_nc_water', y = 'Count', title = paste("Histogram of", 'cova_il_nc_water'))+
  theme_light()

ggplot(dataset, aes(x = .data[['cova_il_nc_water']])) +
  geom_density() + 
  labs(x = 'cova_il_nc_water', y = 'Density', title = paste("Density of", 'cova_il_nc_water'))

  vec <- dataset$cova_il_nc_water
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]

ggplot(dataset, aes(sample = .data[['cova_il_nc_water']], col='red')) + 
  stat_qq() + 
  geom_abline(slope = slope, intercept = int) +
  theme(legend.position = "none") +
  labs(y = 'cova_il_nc_water', title = paste("QQ Plot of", 'cova_il_nc_water'))

ggplot(dataset, aes(x = .data[['log_cova_il_nc_water']])) +
  geom_histogram(bins=100) +
  labs(x = 'log(cova_il_nc_water)', y = 'Count', title = paste("Histogram of log ", 'cova_il_nc_water'))+
  theme_light()

ggplot(dataset, aes(x = .data[['log_cova_il_nc_water']])) +
  geom_density() + 
  labs(x = 'log(cova_il_nc_water)', y = 'Density', title = 'Density of log(cova_il_nc_water)')

  vec <- dataset$log_cova_il_nc_water
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]

ggplot(dataset, aes(sample = .data[['log_cova_il_nc_water']], col='red')) + 
  stat_qq() + 
  geom_abline(slope = slope, intercept = int) +
  theme(legend.position = "none") +
  labs(y = 'log(cova_il_nc_water)', title = 'QQ Plot of log(cova_il_nc_water)')
```

#### Shapiro test


```{r}
shapiro.test(dataset[sample(nrow(dataset), 5000), 'log_cova_il_nc_water'])
```
The p-value > 0.05 implying that the distribution of the data are not significantly different from normal distribution. In other words, we can NOT assume the normality.

```{r}
plot(dataset$log_cova_il_nc_water, pch=20)
```
```{r}
plotdist(dataset$log_cova_il_nc_water, histo = TRUE, demp = TRUE)
```
```{r}
descdist(dataset$log_cova_il_nc_water, discrete=FALSE, boot=500)
```

```{r}
fit_n  <- fitdist(dataset$log_cova_il_nc_water, "norm", method="mle")
```
```{r}
summary(fit_n)
```
```{r}
plot.legend <- c("Normal")
denscomp(list(fit_n), legendtext = plot.legend)
cdfcomp (list(fit_n), legendtext = plot.legend)
```

