---
title: "EDA Severity of Water peril claims in CovA. Visualization"
output:
  html_document:
    df_print: paged
---

```{r}
dataset <- read.csv("C:\\Kate\\Research\\Property\\Data\\property_water_claims_non_cat_fs_v5.csv", header=TRUE)
```

Claim Partial Dependency XGB Classification dataset exported from a Python notebook
```{r}
pd_gamma_dataset <- read.csv("C:\\Kate\\Research\\Property\\Data\\EDA_Severity_Gamma_PartialDependency.csv", header=TRUE)
pd_normal_dataset <- read.csv("C:\\Kate\\Research\\Property\\Data\\EDA _Severity_Normal_PartialDependency.csv", header=TRUE)
```

```{r}
library(dplyr)
library(funModeling)
library(ggplot2)
```

```{r}
colnames(dataset)
```


```{r}
dataset <- dataset[dataset$cova_il_nc_water>0,]
```

## Quick Overview





```{r}
str(dataset)
```

```{r}
summary(dataset)
```

## Visualization

### Severity


```{r}
ggplot(dataset, aes(x = .data[['cova_il_nc_water']], color="red",fill="#56B4E9")) +
  geom_histogram(bins=100) +
  labs(x = 'cova_il_nc_water', y = 'Count', title = "Histogram of Losses (cova_il_nc_water)")+
  scale_color_manual(values=c("red")) +  
  scale_fill_manual(values=c("#56B4E9")) +  
  theme(legend.position = "none")

ggplot(dataset, aes(x = .data[['cova_il_nc_water']])) +
  geom_density() + 
  labs(x = 'cova_il_nc_water', y = 'Density', title = "Density of Losses (cova_il_nc_water)")

  vec <- dataset$cova_il_nc_water
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]

ggplot(dataset, aes(sample = .data[['cova_il_nc_water']], col='red')) + 
  stat_qq() + 
  geom_abline(slope = slope, intercept = int) +
  theme(legend.position = "none") +
  labs(y = 'cova_il_nc_water', title = "QQ Plot of Losses (cova_il_nc_water)")

ggplot(dataset, aes(x = .data[['log_cova_il_nc_water']], color="red",fill="#56B4E9")) +
  geom_histogram(bins=100) +
  scale_color_manual(values=c("red")) +  
  scale_fill_manual(values=c("#56B4E9")) + 
  labs(x = 'log(cova_il_nc_water)', y = 'Count', title = "Histogram of Losses (log_cova_il_nc_water)")+
  theme(legend.position = "none")

ggplot(dataset, aes(x = .data[['log_cova_il_nc_water']])) +
  geom_density() + 
  labs(x = 'log(cova_il_nc_water)', y = 'Density', title = 'Density of Losses (log_cova_il_nc_water)')

  vec <- dataset$log_cova_il_nc_water
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]

ggplot(dataset, aes(sample = .data[['log_cova_il_nc_water']], col='red')) + 
  stat_qq() + 
  geom_abline(slope = slope, intercept = int) +
  theme(legend.position = "none") +
  labs(y = 'log(cova_il_nc_water)', title = 'QQ Plot of Losses (log_cova_il_nc_water)')



```


Looks like gamma distribution and log conversion is more close to normal.

### Box Plots of categorical predictors and Losses
```{r}
plot_categorical <- function(col_name) {
#dataset[[col_name]] <- as.factor(dataset[[col_name]])
#ggplot(dataset, aes(x=.data[[col_name]], y=cova_il_nc_water, col=.data[[col_name]], fill=.data[[col_name]])) +
#  geom_boxplot(notch=notch) +
#  theme(legend.position = "none") +
#  labs(x = col_name, title = paste("Box Plot of Losses and ", col_name))
if ( col_name == 'roofcd_encd') {
col_name_original <- 'roofcd'
} 
  else if ( col_name == 'usagetype_encd' ) {
col_name_original <- 'usagetype'
} 
else if ( col_name == 'constructioncd_encd' ) {
col_name_original <- 'constructioncd'
} 
  else if ( col_name == 'occupancy_encd' ) {
col_name_original <- 'occupancycd'
} 
  else {
  col_name_original <- col_name
}
  
if (grepl('encd',col_name) | 
    col_name =='fire_risk_model_score' | 
    col_name =='customer_cnt_active_policies_binned' |
    col_name =='cova_deductible' | 
    col_name =='cova_limit' | 
    col_name =='protectionclass' | 
    col_name =='ordinanceorlawpct' | 
    col_name =='numberoffamilies' | 
    col_name =='waterded' | 
    col_name =='units' | 
    col_name =='sprinklersystem' |
    col_name =='stories') {
  XBreaks <- as.vector(unlist(unique(dataset[c(col_name)])))
  XLabels <- as.vector(unlist(unique(dataset[c(col_name_original)]))) 
} 
else if (col_name =='pipe_froze_3_blk' | 
col_name =='water_risk_3_blk' | 
col_name =='ustructure_fail_3_blk' | 
col_name =='water_risk_fre_3_blk' | 
col_name =='waterh_fail_3_blk' | 
col_name =='rep_cost_3_blk' | 
col_name =='plumb_leak_3_blk' | 
col_name =='appl_fail_3_blk' | 
col_name =='fixture_leak_3_blk') {
  XBreaks <- c(0,1,2,3,4,5)
  XLabels <- c('Low', 'Elevated', 'Below Avg', 'Average', 'High', 'Highest')    
}
else {
  XBreaks <- c(0,1)
  XLabels <- c("No", "Yes")  

}
  
pd_col <- pd_gamma_dataset[pd_gamma_dataset$feature == col_name,][c('value','pd')]
colnames(pd_col) <- c(col_name,'pd')
plotdata <- dataset %>%
  group_by(.dots = col_name) %>%
  dplyr::summarize(n = n(),
         mean = mean(cova_il_nc_water),
         sd = sd(cova_il_nc_water),
         se = sd / sqrt(n) #,
         #ci = qt(0.975, df = n - 1) * sd / sqrt(n)
         )

plotdata <- merge(plotdata,pd_col,by=col_name)  
cols <- c('Mean Losses with standard error'='#f04546','Losses Partial Dependency'='#3591d1')
p1 = ggplot(plotdata) +
scale_fill_manual(values=cols) +
scale_colour_manual(values=cols) +  
geom_point(aes(x = .data[[col_name]], y = mean,group = 1,color='Mean Losses with standard error'), size = 3) +
geom_line(aes(x = .data[[col_name]], y = mean,group = 1,color='Mean Losses with standard error')) +
geom_errorbar(aes(x = .data[[col_name]], y = mean,group = 1, ymin = mean - se,ymax = mean + se,color='Mean Losses with standard error'),width = .1) +
geom_point(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses Partial Dependency'), size = 3) +  
geom_line(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses Partial Dependency')) +
scale_x_continuous( breaks=XBreaks, labels= XLabels)   +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


p2 = ggplot(plotdata) +
scale_fill_manual(values=cols) +
scale_colour_manual(values=cols) +    
geom_point(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses Partial Dependency'), size = 3) +  
geom_line(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses Partial Dependency')) +
scale_x_continuous( breaks=XBreaks, labels= XLabels)   +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

print(p1)
print(p2)

#----------------------
pd_col <- pd_normal_dataset[pd_normal_dataset$feature == col_name,][c('value','pd')]
if (nrow(pd_col)>0) {
colnames(pd_col) <- c(col_name,'pd')
plotdata <- dataset %>%
  group_by(.dots = col_name) %>%
  dplyr::summarize(n = n(),
         mean = mean(log_cova_il_nc_water),
         sd = sd(log_cova_il_nc_water),
         se = sd / sqrt(n) #,
         #ci = qt(0.975, df = n - 1) * sd / sqrt(n)
         )

plotdata <- merge(plotdata,pd_col,by=col_name)  

cols <- c('Mean Losses (Log) with standard error'='#f04546','Losses (Log) Partial Dependency'='#3591d1')
p1 = ggplot(plotdata) +
scale_fill_manual(values=cols) +
scale_colour_manual(values=cols) +    
geom_point(aes(x = .data[[col_name]], y = mean,group = 1,color='Mean Losses (Log) with standard error'), size = 3) +
geom_line(aes(x = .data[[col_name]], y = mean,group = 1,color='Mean Losses (Log) with standard error')) +
geom_errorbar(aes(x = .data[[col_name]], y = mean,group = 1, ymin = mean - se,ymax = mean + se,color='Mean Losses (Log) with standard error'),width = .1) +
geom_point(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses (Log) Partial Dependency'), size = 3) +  
geom_line(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses (Log) Partial Dependency')) +
scale_x_continuous( breaks=XBreaks, labels= XLabels)   +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


p2 = ggplot(plotdata) +
scale_fill_manual(values=cols) +
scale_colour_manual(values=cols) +    
geom_point(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses (Log) Partial Dependency'), size = 3) +  
geom_line(aes(x = .data[[col_name]], y = pd,group = 1,color='Losses (Log) Partial Dependency')) +
scale_x_continuous( breaks=XBreaks, labels= XLabels)   +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

print(p1)
print(p2)
}
}
```




```{r}
plot_categorical('functionalreplacementcost')
```
```{r}
plot_categorical('numberoffamilies')
```
```{r}
plot_categorical('units')
```
```{r}
plot_categorical('stories')
```
```{r,fig.width=10}
plot_categorical('cova_deductible')
```
```{r}
plot_categorical('replacementvalueind')
```
```{r}
plot_categorical('neighborhoodcrimewatchind')
```
```{r}
plot_categorical('fire_risk_model_score')
```

```{r}
ggplot(data=dataset, mapping = aes(x=water_risk_sev_3_blk, y=cova_il_nc_water, color=-cova_il_nc_water)) +
  geom_point() +
  labs(title = "Losses vs water_risk_sev_3_blk")
```

The higest losses are between 100 and 200 water_risk_sev_3_blk. We may have not enough data for higher numbers of water_risk_sev_3_blk

```{r}
plot_categorical('burglaryalarmtype')
```
```{r}
plot_categorical('deadboltind')
```
```{r}
plot_categorical('waterh_fail_3_blk')
```
```{r}
plot_categorical('safeguardplusind')
```
```{r}
plot_categorical('equipmentbreakdown')
```
```{r}
plot_categorical('cova_limit')
```


```{r}
ggplot(data=dataset, mapping = aes(x=water_risk_3_blk, y=cova_il_nc_water, color=-cova_il_nc_water)) +
  geom_point() +
  labs(title = "Losses vs water_risk_3_blk")
```
There are more higher losses  around 250 water_risk_3_blk

```{r}
plot_categorical('usagetype_encd')
```
```{r}
plot_categorical('multipolicyindumbrella')
```
```{r}
plot_categorical('constructioncd_encd')
```
```{r}
plot_categorical('replacementcostdwellingind')
```
```{r}
plot_categorical('roofcd_encd')
```
```{r}
plot_categorical('gatedcommunityind')
```
```{r}
plot_categorical('homegardcreditind')
```
```{r}
plot_categorical('ustructure_fail_3_blk')
```
```{r}
ggplot(data=dataset, mapping = aes(x=water_risk_fre_3_blk, y=cova_il_nc_water, color=-cova_il_nc_water)) +
  geom_point() +
  labs(title = "Losses vs water_risk_fre_3_blk")
```



```{r}
plot_categorical('ecy')
```
```{r}
ggplot(data=dataset, mapping = aes(x=ecy, y=cova_il_nc_water, color=-cova_il_nc_water)) +
  geom_point() +
  labs(title = "Losses vs ecy")
```

```{r}
plot_categorical('appl_fail_3_blk')
```
```{r}
plot_categorical('pipe_froze_3_blk')
```
```{r}
plot_categorical('firealarmtype')
```
```{r}
plot_categorical('fixture_leak_3_blk')
```
```{r}
plot_categorical('sqft')
```

```{r}
ggplot(data=dataset, mapping = aes(x=sqft, y=cova_il_nc_water, color=-cova_il_nc_water)) +
  geom_point() +
  labs(title = "Losses vs sqft")
```
Severity is higher in larger properties

```{r}
plot_categorical('yearbuilt')
```

```{r}
ggplot(data=dataset, mapping = aes(x=yearbuilt, y=cova_il_nc_water, color=-cova_il_nc_water)) +
  geom_point() +
  labs(title = "Losses vs YearBuilt")
```

Severity is higher in newer properties

```{r}
plot_categorical('customer_cnt_active_policies_binned')
```
```{r}
ggplot(data=dataset, mapping = aes(x=customer_cnt_active_policies_binned, y=cova_il_nc_water, color=-cova_il_nc_water)) +
  geom_point() +
  labs(title = "Losses vs customer_cnt_active_policies_binned")
```



```{r}
plot_categorical('protectionclass')
```
```{r}
plot_categorical('multipolicyind')
```
```{r}
plot_categorical('rep_cost_3_blk')
```
```{r}
plot_categorical('plumb_leak_3_blk')
```
```{r}
plot_categorical('rentersinsurance')
```
```{r}
plot_categorical('sprinklersystem')
```
```{r}
plot_categorical('kitchenfireextinguisherind')
```
```{r}
plot_categorical('poolind')
```
```{r}
plot_categorical('landlordind')
```
```{r}
plot_categorical('occupancy_encd')
```
```{r}
plot_categorical('ordinanceorlawpct')
```
```{r}
plot_categorical('waterded')
```
```{r}
plot_categorical('serviceline')
```
```{r}
plot_categorical('propertymanager')
```
```{r}
plot_categorical('earthquakeumbrellaind')
```








### Misc



```{r}
ggplot(data=dataset, mapping = aes(x=yearbuilt, y=cova_il_nc_water, color=roofcd)) +
  geom_point() +
  labs(title = "Losses vs YearBuilt and Roofcd Limit")
```



```{r}
ggplot(data=dataset, mapping = aes(x=sqft, y=cova_il_nc_water, color=-cova_limit)) +
  geom_point() +
  labs(title = "Losses vs Sqft and CovA Limit")
```